---
title: "WNBA & NBA Statistical Leaders"
subtitle: "Analyzing Playing Time and Box Score Production Across Professional Basketball (2025–2026)"
format:
  html:
    theme: flatly
    page-layout: full
    toc: false
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(plotly)

# WNBA individual game box scores (2025-2026 seasons)
games <- read_csv("data/wnba_player_games.csv", show_col_types = FALSE)

# NBA individual game box scores from ESPN (2025-2026 regular season)
nba_games <- read_csv("data/nba_player_games.csv", show_col_types = FALSE)

# Sample thousands of individual game performances for fast, responsive interactive plots
set.seed(42)
wnba_sample <- games |> slice_sample(n = 2000)
nba_sample <- nba_games |> slice_sample(n = 2000)

# Helper function to generate clean, responsive interactive scatter plots
make_scatter_plotly <- function(sample_data, full_data, metric_col, metric_label, league_label, color_code, count_label) {
  p <- sample_data |>
    mutate(
      hover_text = paste0(
        "<b>", player, "</b> (", team, ")<br>",
        "Date: ", game_date, "<br>",
        "Minutes: ", round(minutes, 1), "<br>",
        metric_label, ": ", .data[[metric_col]]
      )
    ) |>
    ggplot(aes(x = minutes, y = .data[[metric_col]])) +
    geom_point(aes(text = hover_text), alpha = 0.35, size = 1.1, color = color_code) +
    geom_smooth(data = full_data, method = "lm", formula = y ~ x, color = "#2c3e50", linetype = "dashed", linewidth = 0.9, se = FALSE) +
    labs(
      title = paste0(league_label, ": ", metric_label, " vs. Minutes"),
      subtitle = count_label,
      x = "Minutes Played",
      y = metric_label
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 13, color = "#2c3e50"),
      panel.grid.minor = element_blank(),
      plot.margin = margin(10, 15, 10, 15)
    )
  
  ggplotly(p, tooltip = "text", height = 390) |>
    config(displaylogo = FALSE, modeBarButtonsToRemove = c("lasso2d", "select2d"))
}
```

::: {.panel-tabset}

## Exploration

::: {.lead style="color: #5a6268; margin-bottom: 20px;"}
Explore how playing time influences production across both leagues. Use the filter below to switch categories—the **WNBA** chart appears first, with the corresponding **NBA** chart directly underneath.
:::

```{=html}
<!-- Control Panel Card -->
<div class="card shadow-sm border-0 mb-4" style="background: linear-gradient(135deg, #f8f9fa 0%, #eef2f7 100%); border-radius: 12px; border-left: 5px solid #3498db !important;">
  <div class="card-body d-flex flex-wrap align-items-center justify-content-between p-3">
    <div class="d-flex align-items-center mb-2 mb-md-0">
      <div style="font-size: 1.8rem; margin-right: 14px; line-height: 1;">📊</div>
      <div>
        <div class="fw-bold" style="font-size: 1.1rem; color: #2c3e50;">Select Statistical Category</div>
        <small class="text-muted">Both charts will dynamically update to display the chosen metric</small>
      </div>
    </div>
    <div class="d-flex align-items-center">
      <select id="metric-select" class="form-select form-select-lg shadow-sm" style="min-width: 230px; font-weight: 600; font-size: 15px; border-radius: 8px; border: 1px solid #ced4da; cursor: pointer; background-color: #ffffff;">
        <option value="points" selected>🏀 Points</option>
        <option value="rebounds">🔄 Rebounds</option>
        <option value="blocks">🛡️ Blocks</option>
        <option value="assists">🤝 Assists</option>
        <option value="steals">⚡ Steals</option>
        <option value="fouls">⚠️ Personal Fouls</option>
      </select>
    </div>
  </div>
</div>
```

<!-- POINTS SECTION -->
::: {#stat-pair-points .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #e66101; font-size: 1.05rem;">WNBA · Points Scored</span>
    <span class="badge" style="background-color: #e66101; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-points
#| warning: false
make_scatter_plotly(wnba_sample, games, "points", "Points Scored", "WNBA", "#e66101", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #2980b9; font-size: 1.05rem;">NBA · Points Scored</span>
    <span class="badge" style="background-color: #2980b9; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-points
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "points", "Points Scored", "NBA", "#2980b9", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

<!-- REBOUNDS SECTION -->
::: {#stat-pair-rebounds .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #2ca25f; font-size: 1.05rem;">WNBA · Rebounds Grabbed</span>
    <span class="badge" style="background-color: #2ca25f; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-rebounds
#| warning: false
make_scatter_plotly(wnba_sample, games, "rebounds", "Rebounds Grabbed", "WNBA", "#2ca25f", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #2ca25f; font-size: 1.05rem;">NBA · Rebounds Grabbed</span>
    <span class="badge" style="background-color: #2ca25f; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-rebounds
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "rebounds", "Rebounds Grabbed", "NBA", "#2ca25f", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

<!-- BLOCKS SECTION -->
::: {#stat-pair-blocks .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #ca0020; font-size: 1.05rem;">WNBA · Blocks</span>
    <span class="badge" style="background-color: #ca0020; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-blocks
#| warning: false
make_scatter_plotly(wnba_sample, games, "blocks", "Blocks", "WNBA", "#ca0020", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #ca0020; font-size: 1.05rem;">NBA · Blocks</span>
    <span class="badge" style="background-color: #ca0020; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-blocks
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "blocks", "Blocks", "NBA", "#ca0020", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

<!-- ASSISTS SECTION -->
::: {#stat-pair-assists .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #2b83ba; font-size: 1.05rem;">WNBA · Assists</span>
    <span class="badge" style="background-color: #2b83ba; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-assists
#| warning: false
make_scatter_plotly(wnba_sample, games, "assists", "Assists", "WNBA", "#2b83ba", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #2b83ba; font-size: 1.05rem;">NBA · Assists</span>
    <span class="badge" style="background-color: #2b83ba; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-assists
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "assists", "Assists", "NBA", "#2b83ba", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

<!-- STEALS SECTION -->
::: {#stat-pair-steals .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #7b3294; font-size: 1.05rem;">WNBA · Steals</span>
    <span class="badge" style="background-color: #7b3294; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-steals
#| warning: false
make_scatter_plotly(wnba_sample, games, "steals", "Steals", "WNBA", "#7b3294", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #7b3294; font-size: 1.05rem;">NBA · Steals</span>
    <span class="badge" style="background-color: #7b3294; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-steals
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "steals", "Steals", "NBA", "#7b3294", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

<!-- PERSONAL FOULS SECTION -->
::: {#stat-pair-fouls .metric-stat-pair}
```{=html}
<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #5e3c99; font-size: 1.05rem;">WNBA · Personal Fouls</span>
    <span class="badge" style="background-color: #5e3c99; color: white;">WNBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-wnba-fouls
#| warning: false
make_scatter_plotly(wnba_sample, games, "fouls", "Personal Fouls", "WNBA", "#5e3c99", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>

<div class="card shadow-sm border-0 mb-4 plot-card">
  <div class="card-header bg-white d-flex justify-content-between align-items-center py-2 px-3 border-bottom">
    <span class="fw-bold" style="color: #5e3c99; font-size: 1.05rem;">NBA · Personal Fouls</span>
    <span class="badge" style="background-color: #5e3c99; color: white;">NBA (2025–2026)</span>
  </div>
  <div class="card-body p-2">
```
```{r}
#| label: graph-nba-fouls
#| warning: false
make_scatter_plotly(nba_sample, nba_games, "fouls", "Personal Fouls", "NBA", "#5e3c99", "2,000 sampled game performances (2025–2026)")
```
```{=html}
  </div>
</div>
```
:::

```{=html}
<!-- Footer / Methodology Card -->
<div class="card bg-light border-0 mt-4 mb-5" style="border-radius: 10px;">
  <div class="card-body text-muted small p-3">
    <strong>💡 Interactive Tips & Notes:</strong>
    <ul class="mb-1 mt-1">
      <li><strong>Hover</strong> over individual points to inspect the player name, team, date, minutes, and exact statistics.</li>
      <li><strong>Click and drag</strong> across any plot to zoom into clusters; double-click to reset the axes.</li>
      <li><strong>Dashed trendline</strong> represents an ordinary least squares (OLS) linear regression model fitted across all 12,013 individual player performances.</li>
      <li><strong>Data Sources:</strong> SportsDataverse (<code>wehoop</code> for WNBA and <code>hoopR</code> via ESPN for NBA).</li>
    </ul>
  </div>
</div>

<style>
.metric-stat-pair {
  display: none;
  width: 100%;
}
.metric-stat-pair.active {
  display: block;
}
.plot-card {
  border-radius: 12px;
  overflow: hidden;
  border: 1px solid #eef2f5 !important;
  transition: transform 0.15s ease-in-out, box-shadow 0.15s ease-in-out;
}
.plot-card:hover {
  box-shadow: 0 6px 18px rgba(0, 0, 0, 0.08) !important;
}
</style>

<script>
function updateMetricPair() {
  const metricSelect = document.getElementById("metric-select");
  if (!metricSelect) return;

  const metric = metricSelect.value;
  const targetId = "stat-pair-" + metric;

  document.querySelectorAll(".metric-stat-pair").forEach(function(el) {
    el.style.display = "none";
    el.classList.remove("active");
  });

  const target = document.getElementById(targetId);
  if (target) {
    target.style.display = "block";
    target.classList.add("active");

    // Resize plotly plots inside the active container
    target.querySelectorAll(".js-plotly-plot").forEach(function(plotEl) {
      if (window.Plotly) {
        window.Plotly.Plots.resize(plotEl);
      }
    });
    window.dispatchEvent(new Event("resize"));
  }
}

function initMetricDropdown() {
  const metricSelect = document.getElementById("metric-select");
  if (metricSelect) {
    metricSelect.addEventListener("change", updateMetricPair);
    updateMetricPair();
  }
}

if (document.readyState === "loading") {
  document.addEventListener("DOMContentLoaded", initMetricDropdown);
} else {
  initMetricDropdown();
}

document.addEventListener("shown.bs.tab", function() {
  updateMetricPair();
});
</script>
```

## About

### Project Overview

This project explores the relationship between playing time (minutes) and statistical output across both the **WNBA** and **NBA** for the 2025–2026 seasons.

- **Data Sources:** 
  - WNBA: SportsDataverse (`wehoop`)
  - NBA: ESPN (`hoopR`)
- **Key Categories:** Points, Rebounds, Assists, Steals, Blocks, and Personal Fouls.
- **Sample:** Both leagues display 2,000 sampled individual game performances with linear regression trendlines calculated across all 12,013 games.

### About



:::

