---
title: "WNBA & NBA Leaders"
format:
  html:
    page-layout: full
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(plotly)

# WNBA individual game box scores (2025-2026 seasons)
games <- read_csv("data/wnba_player_games.csv", show_col_types = FALSE)

# NBA individual game box scores from ESPN (2025-2026 regular season)
nba_games <- read_csv("data/nba_player_games.csv", show_col_types = FALSE)

# Sample thousands of individual game performances for fast, responsive interactive plots
set.seed(42)
wnba_sample <- games |> slice_sample(n = 2000)
nba_sample <- nba_games |> slice_sample(n = 2000)

# Helper function to generate clean, responsive interactive scatter plots
make_scatter_plotly <- function(sample_data, full_data, metric_col, metric_label, league_label, color_code, count_label) {
  p <- sample_data |>
    mutate(
      hover_text = paste0(
        "<b>", player, "</b> (", team, ")<br>",
        "Date: ", game_date, "<br>",
        "Minutes: ", round(minutes, 1), "<br>",
        metric_label, ": ", .data[[metric_col]]
      )
    ) |>
    ggplot(aes(x = minutes, y = .data[[metric_col]])) +
    geom_point(aes(text = hover_text), alpha = 0.35, size = 1.1, color = color_code) +
    geom_smooth(data = full_data, method = "lm", formula = y ~ x, color = "black", linetype = "dashed", se = FALSE) +
    labs(
      title = paste0(league_label, ": ", metric_label, " vs. Minutes<br><sup>", count_label, "</sup>"),
      x = "Minutes Played",
      y = metric_label
    ) +
    theme_minimal(base_size = 11)
  
  ggplotly(p, tooltip = "text", height = 400)
}
```

## Exploring Playing Time and Performance in the WNBA & NBA

How does playing time impact individual player output on the court across both leagues? We examine the relationships between minutes played and key statistical categories. 

Each plot contains thousands of individual player-game performances from the 2025–2026 seasons (with the dashed trend line fitted across all games). **Hover** over any point to inspect the player, date, minutes, and stat line, or **click and drag** to zoom into any cluster!

### Points vs. Minutes

Scoring opportunity scales strongly with court time in both leagues.

::: {layout-ncol=2}
```{r}
#| label: points-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "points", "Points Scored", "WNBA", "#e66101", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: points-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "points", "Points Scored", "NBA", "#e66101", "2,000 sampled game performances (2025–2026)")
```
:::

### Rebounds vs. Minutes

Rebounding shows a steady positive slope with increased playing time in both leagues.

::: {layout-ncol=2}
```{r}
#| label: rebounds-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "rebounds", "Rebounds Grabbed", "WNBA", "#2ca25f", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: rebounds-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "rebounds", "Rebounds Grabbed", "NBA", "#2ca25f", "2,000 sampled game performances (2025–2026)")
```
:::

### Assists vs. Minutes

Playmaking is heavily correlated with court time across both leagues, especially for starting guards.

::: {layout-ncol=2}
```{r}
#| label: assists-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "assists", "Assists", "WNBA", "#2b83ba", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: assists-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "assists", "Assists", "NBA", "#2b83ba", "2,000 sampled game performances (2025–2026)")
```
:::

### Steals vs. Minutes

Steals occur less frequently per game than points or rebounds, but extended minutes heighten takeaways in both leagues.

::: {layout-ncol=2}
```{r}
#| label: steals-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "steals", "Steals", "WNBA", "#7b3294", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: steals-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "steals", "Steals", "NBA", "#7b3294", "2,000 sampled game performances (2025–2026)")
```
:::

### Blocks vs. Minutes

Shot blocking shows a gradual positive slope with minutes, driven primarily by rim protectors.

::: {layout-ncol=2}
```{r}
#| label: blocks-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "blocks", "Blocks", "WNBA", "#ca0020", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: blocks-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "blocks", "Blocks", "NBA", "#ca0020", "2,000 sampled game performances (2025–2026)")
```
:::

### Personal Fouls vs. Minutes

Defending for longer stretches raises the likelihood of personal fouls in both leagues.

::: {layout-ncol=2}
```{r}
#| label: fouls-vs-minutes-wnba
#| warning: false

make_scatter_plotly(wnba_sample, games, "fouls", "Personal Fouls", "WNBA", "#5e3c99", "2,000 sampled game performances (2025–2026)")
```

```{r}
#| label: fouls-vs-minutes-nba
#| warning: false

make_scatter_plotly(nba_sample, nba_games, "fouls", "Personal Fouls", "NBA", "#5e3c99", "2,000 sampled game performances (2025–2026)")
```
:::
